Finite dimensional Lie algebras

### 2010-03-15 AffineLieAlgebras/SchemePrograms/LiterateLie.tex

\documentclass[a4paper,10pt]{article}% ===> this file was generated automatically by noweave --- better not edit it
\usepackage[unicode,verbose]{hyperref}
\usepackage{amsmath,amssymb,amsthm} \usepackage{pb-diagram}
\usepackage{ucs}
\usepackage{noweb}
%\usepackage[utf8x]{inputenc}
%\usepackage[russian]{babel}
\usepackage{cmap}
\usepackage[pdftex]{graphicx}
\pagestyle{plain}
\theoremstyle{definition} \newtheorem{Def}{Definition}
%\usepackage{verbatim} 
\newenvironment{comment}
{\par\noindent{\bf TODO}\\}
{\\\hfill$\scriptstyle\blacksquare$\par}

@@

We will represent Lie algebra in the code as the object of class {\Tt{}lie-algebra\nwendquote}. We use very simple class system, which is described in the Appendix \ref{sec:object-system}. So {\Tt{}lie-algebra\nwendquote} is really an interface, but since our object system doesn't have interfaces we make it an abstract class and concrete classes such as {\Tt{}semisimple-lie-algebra\nwendquote} inherit it.
Lie algebra is a vector space, so we really should create class for the vector space which contains methods for the change of basis. But for now we use only one standard basis for Lie algebras and {\Tt{}lie-algebra\nwendquote} is the subclass of the most genral class {\Tt{}object\nwendquote}.
\nwenddocs{}\nwbegincode{2}\sublabel{NW1VK18x-4dzmez-1}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-4dzmez-1}}}\moddef{Lie algebra class .scm~{\nwtagstyle{}\subpageref{NW1VK18x-4dzmez-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-3yKaW-1}}\nwenddeflinemarkup
(class 'lie-algebra 'object
       \LA{}Lie algebra methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-4X6QTQ-1}}\RA{})
\nwused{\\{NW1VK18x-3yKaW-1}}\nwendcode{}\nwbegindocs{3}\nwdocspar
We need to add several definitions and describe a structure of Lie algebra in order to discuss fields and methods of {\Tt{}lie-algebra\nwendquote} class.
\begin{Def}
  {\bf Ideal} $\mathfrak{e}\subset\mathfrak{g}$: $[\mathfrak{g},\mathfrak{e}]\subset \mathfrak{e}$
\end{Def}

@@

\begin{Def}
  {\bf Cartan subalgebra $\mathfrak{h}\subset\mathfrak{g}$} - is the maximal commutative subalgebra of $\mathfrak{g}$. The dimension of the Cartan subalgebra $\mathfrak{h}$ is called {\bf rank} of Lie algebra $\mathfrak{g}$.
\end{Def}
\nwenddocs{}\nwbegincode{4}\sublabel{NW1VK18x-4X6QTQ-1}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-4X6QTQ-1}}}\moddef{Lie algebra methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-4X6QTQ-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-4dzmez-1}}\nwenddeflinemarkup
`(rank ,(lambda (self)
              (error "Abstract class!")))
`(dimension ,(lambda (self)
              (error "Abstract class!")))
`(cartan-subalgebra ,(lambda (self) 
              (error "Abstract class!")))

\nwused{\\{NW1VK18x-4dzmez-1}}\nwendcode{}\nwbegindocs{5}\nwdocspar
Commutative algebras constitute first concrete class of Lie algebras.
\nwenddocs{}\nwbegincode{6}\sublabel{NW1VK18x-I1QOi-1}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-I1QOi-1}}}\moddef{Commutative algebra .scm~{\nwtagstyle{}\subpageref{NW1VK18x-I1QOi-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-3yKaW-1}}\nwenddeflinemarkup
(class 'commutative-algebra 'lie-algebra
       '(dim 1)
       `(rank ,(lambda (self) (send self 'dim)))
       `(dimension ,(lambda (self) (send self 'dim)))
       `(cartan-subalgebra ,(lambda (self) self)))

;; The constructor
(define (make-commutative-algebra dim)
  (new 'commutative-algebra `(dim ,dim)))
\nwused{\\{NW1VK18x-3yKaW-1}}\nwendcode{}\nwbegindocs{7}\nwdocspar
\begin{comment}
It can be interesting and useful to add a method which show commutation relatons of the algebra in the symbolic form, but this task is left for the future  
\end{comment}
\nwenddocs{}\nwbegindocs{8}\nwdocspar

\section{Semisimple Lie algebras}
\label{sec:semis-lie-algebra}
The structure of simple and semisimple Lie algebras can be encoded by the set of simple roots. 
\nwenddocs{}\nwbegincode{9}\sublabel{NW1VK18x-1QMleH-1}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-1QMleH-1}}}\moddef{Semisimple Lie algebra .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1QMleH-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-3yKaW-1}}\nwenddeflinemarkup
(class 'semisimple-lie-algebra 'lie-algebra
       '(simple-roots ())
       \LA{}Semisimple Lie algebra methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1pDr9o-1}}\RA{})
;; The constructor
(define (make-semisimple-lie-algebra simple-roots)
  (new 'semisimple-lie-algebra `(simple-roots ,simple-roots)))

\nwused{\\{NW1VK18x-3yKaW-1}}\nwendcode{}\nwbegindocs{10}\nwdocspar

\subsection{Root systems}
\label{sec:root-systems}
\begin{Def}
  In the {\bf Cartan-Weyl} basis the genrators are constructed as follows.
  First we choose the genrators $H_i\in \mathfrak{h}$ of the Cartan subalgebra. All these generators commute $[H_i,H_j]=0$ and their representations can be diagonalised simultaneously. The remaining generators $E_{\alpha}$ are chosen to be a linear combinations of $J_a$ such that
  \begin{equation}
    \label{eq:4}
    [H_i,E_{\alpha}]=\alpha_i E_{\alpha}
  \end{equation}
  The vector $\alpha=(\alpha_1,\dots,\alpha_r),\; r=rank(\mathfrak{g})$ is called a {\bf root} and the corresponding operator $E_{\alpha}$ is called {\bf ladder operator}.
\end{Def}
The root $\alpha$ maps $H_i$ to the number $\alpha(H_i)=\alpha_i$ hence $\alpha\in \mathfrak{h}^{*}$ - element of the dual to the Cartan subalgebra. The set of all the roos $\alpha$ is denoted by $\Delta=\{\alpha,\; \alpha \text{ - root}\}$. So it is natural for the {\Tt{}lie-algebra\nwendquote} class to have a method which lists all the roots.
\nwenddocs{}\nwbegincode{11}\sublabel{NW1VK18x-1pDr9o-1}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-1pDr9o-1}}}\moddef{Semisimple Lie algebra methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1pDr9o-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-1QMleH-1}}\nwenddeflinemarkup
`(rank ,(lambda (self)
    (length (send self 'simple-roots))))

`(cartan-subalgebra ,(lambda (self)
    (make-commutative-algebra (send self 'rank))))
\LA{}Root system methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-1}}\RA{}
\nwused{\\{NW1VK18x-1QMleH-1}}\nwendcode{}\nwbegindocs{12}\nwdocspar
All the roots are the linear combinations of simple roots with the integral coefficients. There exists discrete group of symmetry called {\bf Weyl group} which can be used to produce all the roots from the set of simple roots. 
\begin{Def}
  \label{weyl-group-def}
  {\bf Weyl group $W$} of finite-dimension Lie algebra is a finite reflection group of $\mathfrak{h}^*$ generated by the basic reflections corresponding to the simple roots. 
  Each simple root $\alpha_i$ corresponds to the reflection $r_{i}$ in hyperplane orthogonal to $\alpha$.
  \begin{equation}
    \label{eq:5}
    r_{i}\lambda=\lambda-\frac{(\lambda,\alpha_i)}{(\alpha_i,\alpha_i)}\alpha_i
  \end{equation}
  For $w\in W$ there exists several equivalent realizations of the form $w=r_{i_1}\cdot r_{i_2}\dots r_{i_k},\; i_1\dots i_k=1..r$. The expression of the smallest {\it length} is called {\it reduced}. 
\end{Def}
We will implement represent Weyl group elements as the lists of numbers of basic reflection
\begin{equation}
  \label{eq:6}
  w=r_{i_1}\cdot r_{i_2}\dots r_{i_k} \leftrightarrow (i_1, i_2,\dots,i_k)
\end{equation}
Then we will implement action of Weyl group elements on $\mathfrak{h}^*$ using the recipies from \cite{stembridge2001computational}, \cite{casselman1994machine}, but it will be done in the future. For now we are not implementing abstract construction of Weyl group since it is enough to be able to construct Weyl group orbits of $\mathfrak{h}^*$ elements.

So we need just two following general functions:
\nwenddocs{}\nwbegincode{13}\sublabel{NW1VK18x-10zTlp-1}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-10zTlp-1}}}\moddef{Functions for root systems .scm~{\nwtagstyle{}\subpageref{NW1VK18x-10zTlp-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-3yKaW-1}}\nwenddeflinemarkup
(define (co-root r)
  (div (mul 2 r) (prod r r)))

(define (reflect weight root)
  (sub weight (mul (prod weight (co-root root)) root)))
\nwused{\\{NW1VK18x-3yKaW-1}}\nwendcode{}\nwbegindocs{14}\nwdocspar
Here {\it co-root} $\alpha^{\vee}$ is the element of the dual space $(\mathfrak{h}^*)^*$ which is identified with $\mathfrak{h}^*\approx \mathfrak{h}$ since we have scalar product on $\mathfrak{h}$.

Now we can implement wide range of root system related methods of {\Tt{}semisimple-lie-algebra\nwendquote} class
\nwenddocs{}\nwbegincode{15}\sublabel{NW1VK18x-1H68wy-1}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-1}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-1pDr9o-1}}\nwprevnextdefs{\relax}{NW1VK18x-1H68wy-2}\nwenddeflinemarkup
`(simple-co-roots ,(lambda (self)
              (map co-root (send self 'simple-roots))))
\nwalsodefined{\\{NW1VK18x-1H68wy-2}\\{NW1VK18x-1H68wy-3}\\{NW1VK18x-1H68wy-4}\\{NW1VK18x-1H68wy-5}\\{NW1VK18x-1H68wy-6}}\nwused{\\{NW1VK18x-1pDr9o-1}}\nwendcode{}\nwbegindocs{16}\nwdocspar
\begin{Def}
  {\bf Cartan matrix} with the elements
  \begin{equation}
    \label{eq:7}
    a_{ij}=(\alpha_i,\alpha_j^{\vee})=\frac{2(\alpha_i,\alpha_j)}{(\alpha_j,\alpha_j)}
  \end{equation}
  completely determines the set of simple roots and is useful for the classification and compact description of Lie algebras.
\end{Def}
\nwenddocs{}\nwbegincode{17}\sublabel{NW1VK18x-1H68wy-2}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-2}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-1}}}\plusendmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-1pDr9o-1}}\nwprevnextdefs{NW1VK18x-1H68wy-1}{NW1VK18x-1H68wy-3}\nwenddeflinemarkup
`(cartan-matrix
  ,(lambda (self)
     (map (lambda (a)
            (map (lambda (av) (prod a av)) (send self 'simple-co-roots)))
         (send self 'simple-roots))))
\nwused{\\{NW1VK18x-1pDr9o-1}}\nwendcode{}\nwbegindocs{18}\nwdocspar
We have already introduced Weyl group of reflections \ref{weyl-group-def}, now we implement procedure which constructs the union of Weyl group orbits of the set of weights.
\nwenddocs{}\nwbegincode{19}\sublabel{NW1VK18x-1H68wy-3}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-3}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-1}}}\plusendmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-1pDr9o-1}}\nwprevnextdefs{NW1VK18x-1H68wy-2}{NW1VK18x-1H68wy-4}\nwenddeflinemarkup
`(orbit ,
  (lambda (self weights)
    (let ((addon 
           (filter (lambda (x) (not (element-of-set? x weights)))
                   (fold-right union-set '() 
                               (map (lambda (w)
                                      (map
                                       (lambda (x) (reflect w x))
                                       (send self 'simple-roots)))
                                    weights)))))
      (if (null? addon)
          weights
          (send self 'orbit (union-set weights addon) )))))
\nwused{\\{NW1VK18x-1pDr9o-1}}\nwendcode{}\nwbegindocs{20}\nwdocspar
\begin{Def}
  {\bf Fundamental weights} form a basis $w_1,\dots,w_r$ dual to the basis of simple co-roots $\alpha_1^{\vee},\dots,\alpha_r^{\vee}$.
\end{Def}
\nwenddocs{}\nwbegincode{21}\sublabel{NW1VK18x-1H68wy-4}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-4}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-1}}}\plusendmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-1pDr9o-1}}\nwprevnextdefs{NW1VK18x-1H68wy-3}{NW1VK18x-1H68wy-5}\nwenddeflinemarkup
`(fundamental-weights
  ,(lambda (self)
     (map (lambda (x)
            (sum (map-n (lambda (y z) (mul y z)) x (send self 'simple-roots))))
          (matrix-inverse (send self 'cartan-matrix)))))
\nwused{\\{NW1VK18x-1pDr9o-1}}\nwendcode{}\nwbegindocs{22}\nwdocspar
{\bf Weyl vector} $\rho=\sum_i w_i=\frac{1}{2}\sum_{\alpha\in \Delta^{+}}\alpha$.
\nwenddocs{}\nwbegincode{23}\sublabel{NW1VK18x-1H68wy-5}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-5}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-1}}}\plusendmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-1pDr9o-1}}\nwprevnextdefs{NW1VK18x-1H68wy-4}{NW1VK18x-1H68wy-6}\nwenddeflinemarkup
`(rho ,(lambda (self)
         (sum (send self 'fundamental-weights))))

\nwused{\\{NW1VK18x-1pDr9o-1}}\nwendcode{}\nwbegindocs{24}\nwdocspar
We can construct the full set of Lie algebra roots using Weyl symmetry.
\nwenddocs{}\nwbegincode{25}\sublabel{NW1VK18x-1H68wy-6}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-6}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NW1VK18x-1H68wy-1}}}\plusendmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-1pDr9o-1}}\nwprevnextdefs{NW1VK18x-1H68wy-5}{\relax}\nwenddeflinemarkup
`(roots ,(lambda (self)
           (send self 'orbit (send self 'simple-roots))))
\nwused{\\{NW1VK18x-1pDr9o-1}}\nwendcode{}\nwbegindocs{26}\nwdocspar

@@

For the finite-dimensional representations we introduce abstract class {\Tt{}representation\nwendquote} and concrete class {\Tt{}highest-weight-representation\nwendquote}. Then we can describe  tensor product of two highest weight representations as the object of class {\Tt{}representation\nwendquote} (but not of class {\Tt{}highest-weight-representation\nwendquote}.
\nwenddocs{}\nwbegincode{29}\sublabel{NW1VK18x-2w9Axe-1}\nwmargintag{{\nwtagstyle{}\subpageref{NW1VK18x-2w9Axe-1}}}\moddef{Representations .scm~{\nwtagstyle{}\subpageref{NW1VK18x-2w9Axe-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NW1VK18x-3yKaW-1}}\nwprevnextdefs{\relax}{NW1VK18x-2w9Axe-2}\nwenddeflinemarkup
(class 'representation 'object
       `(dim ())
       `(multiplicity , (lambda (self weight)
                          (error "Class is abstract!")))
       '(lie-algebra ())
       )
\nwalsodefined{\\{NW1VK18x-2w9Axe-2}}\nwused{\\{NW1VK18x-3yKaW-1}}\nwendcode{}\nwbegindocs{30}\nwdocspar

>>>>

\documentclass[a4paper,10pt]{article}
\usepackage[unicode,verbose]{hyperref}
\usepackage{amsmath,amssymb,amsthm} \usepackage{pb-diagram}
\usepackage{ucs}
\usepackage{noweb}
%\usepackage[utf8x]{inputenc}
%\usepackage[russian]{babel}
\usepackage{cmap}
\usepackage[pdftex]{graphicx}
\pagestyle{plain}
\theoremstyle{definition} \newtheorem{Def}{Definition}
%\usepackage{verbatim} 
\newenvironment{comment}
{\par\noindent{\bf TODO}\\}
{\\\hfill$\scriptstyle\blacksquare$\par}

@@

We will represent Lie algebra in the code as the object of class {\tt{}lie-algebra}. We use very simple class system, which is described in the Appendix \ref{sec:object-system}. So {\tt{}lie-algebra} is really an interface, but since our object system doesn't have interfaces we make it an abstract class and concrete classes such as {\tt{}semisimple-lie-algebra} inherit it.
Lie algebra is a vector space, so we really should create class for the vector space which contains methods for the change of basis. But for now we use only one standard basis for Lie algebras and {\tt{}lie-algebra} is the subclass of the most genral class {\tt{}object}.
\nwenddocs{}\nwbegincode{2}\sublabel{NWLitE-LieM-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-LieM-1}}}\moddef{Lie algebra class .scm~{\nwtagstyle{}\subpageref{NWLitE-LieM-1}}}\endmoddef
(class 'lie-algebra 'object
       \LA{}Lie algebra methods .scm~{\nwtagstyle{}\subpageref{NWLitE-LieO-1}}\RA{})
\nwused{\\{NWLitE-lieF-1}}\nwendcode{}\nwbegindocs{3}\nwdocspar
We need to add several definitions and describe a structure of Lie algebra in order to discuss fields and methods of {\tt{}lie-algebra} class.
\begin{Def}
  {\bf Ideal} $\mathfrak{e}\subset\mathfrak{g}$: $[\mathfrak{g},\mathfrak{e}]\subset \mathfrak{e}$
\end{Def}

@@

\begin{Def}
  {\bf Cartan subalgebra $\mathfrak{h}\subset\mathfrak{g}$} - is the maximal commutative subalgebra of $\mathfrak{g}$. The dimension of the Cartan subalgebra $\mathfrak{h}$ is called {\bf rank} of Lie algebra $\mathfrak{g}$.
\end{Def}
\nwenddocs{}\nwbegincode{4}\sublabel{NWLitE-LieO-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-LieO-1}}}\moddef{Lie algebra methods .scm~{\nwtagstyle{}\subpageref{NWLitE-LieO-1}}}\endmoddef
`(rank ,(lambda (self)
              (error "Abstract class!")))
`(dimension ,(lambda (self)
              (error "Abstract class!")))
`(cartan-subalgebra ,(lambda (self) 
              (error "Abstract class!")))

\nwused{\\{NWLitE-LieM-1}}\nwendcode{}\nwbegindocs{5}\nwdocspar
Commutative algebras constitute first concrete class of Lie algebras.
\nwenddocs{}\nwbegincode{6}\sublabel{NWLitE-ComO-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-ComO-1}}}\moddef{Commutative algebra .scm~{\nwtagstyle{}\subpageref{NWLitE-ComO-1}}}\endmoddef
(class 'commutative-algebra 'lie-algebra
       '(dim 1)
       `(rank ,(lambda (self) (send self 'dim)))
       `(dimension ,(lambda (self) (send self 'dim)))
       `(cartan-subalgebra ,(lambda (self) self)))

;; The constructor
(define (make-commutative-algebra dim)
  (new 'commutative-algebra `(dim ,dim)))
\nwused{\\{NWLitE-lieF-1}}\nwendcode{}\nwbegindocs{7}\nwdocspar
\begin{comment}
It can be interesting and useful to add a method which show commutation relatons of the algebra in the symbolic form, but this task is left for the future  
\end{comment}
\nwenddocs{}\nwbegindocs{8}\nwdocspar

\section{Semisimple Lie algebras}
\label{sec:semis-lie-algebra}
The structure of simple and semisimple Lie algebras can be encoded by the set of simple roots. 
\nwenddocs{}\nwbegincode{9}\sublabel{NWLitE-SemR-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-SemR-1}}}\moddef{Semisimple Lie algebra .scm~{\nwtagstyle{}\subpageref{NWLitE-SemR-1}}}\endmoddef
(class 'semisimple-lie-algebra 'lie-algebra
       '(simple-roots ())
       \LA{}Semisimple Lie algebra methods .scm~{\nwtagstyle{}\subpageref{NWLitE-SemZ-1}}\RA{})
;; The constructor
(define (make-semisimple-lie-algebra simple-roots)
  (new 'semisimple-lie-algebra `(simple-roots ,simple-roots)))

\nwused{\\{NWLitE-lieF-1}}\nwendcode{}\nwbegindocs{10}\nwdocspar

\subsection{Root systems}
\label{sec:root-systems}
\begin{Def}
  In the {\bf Cartan-Weyl} basis the genrators are constructed as follows.
  First we choose the genrators $H_i\in \mathfrak{h}$ of the Cartan subalgebra. All these generators commute $[H_i,H_j]=0$ and their representations can be diagonalised simultaneously. The remaining generators $E_{\alpha}$ are chosen to be a linear combinations of $J_a$ such that
  \begin{equation}
    \label{eq:4}
    [H_i,E_{\alpha}]=\alpha_i E_{\alpha}
  \end{equation}
  The vector $\alpha=(\alpha_1,\dots,\alpha_r),\; r=rank(\mathfrak{g})$ is called a {\bf root} and the corresponding operator $E_{\alpha}$ is called {\bf ladder operator}.
\end{Def}
The root $\alpha$ maps $H_i$ to the number $\alpha(H_i)=\alpha_i$ hence $\alpha\in \mathfrak{h}^{*}$ - element of the dual to the Cartan subalgebra. The set of all the roos $\alpha$ is denoted by $\Delta=\{\alpha,\; \alpha \text{ - root}\}$. So it is natural for the {\tt{}lie-algebra} class to have a method which lists all the roots.
\nwenddocs{}\nwbegincode{11}\sublabel{NWLitE-SemZ-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-SemZ-1}}}\moddef{Semisimple Lie algebra methods .scm~{\nwtagstyle{}\subpageref{NWLitE-SemZ-1}}}\endmoddef
`(rank ,(lambda (self)
    (length (send self 'simple-roots))))

`(cartan-subalgebra ,(lambda (self)
    (make-commutative-algebra (send self 'rank))))
\LA{}Root system methods .scm~{\nwtagstyle{}\subpageref{NWLitE-RooO-1}}\RA{}
\nwused{\\{NWLitE-SemR-1}}\nwendcode{}\nwbegindocs{12}\nwdocspar
All the roots are the linear combinations of simple roots with the integral coefficients. There exists discrete group of symmetry called {\bf Weyl group} which can be used to produce all the roots from the set of simple roots. 
\begin{Def}
  \label{weyl-group-def}
  {\bf Weyl group $W$} of finite-dimension Lie algebra is a finite reflection group of $\mathfrak{h}^*$ generated by the basic reflections corresponding to the simple roots. 
  Each simple root $\alpha_i$ corresponds to the reflection $r_{i}$ in hyperplane orthogonal to $\alpha$.
  \begin{equation}
    \label{eq:5}
    r_{i}\lambda=\lambda-\frac{(\lambda,\alpha_i)}{(\alpha_i,\alpha_i)}\alpha_i
  \end{equation}
  For $w\in W$ there exists several equivalent realizations of the form $w=r_{i_1}\cdot r_{i_2}\dots r_{i_k},\; i_1\dots i_k=1..r$. The expression of the smallest {\it length} is called {\it reduced}. 
\end{Def}
We will implement represent Weyl group elements as the lists of numbers of basic reflection
\begin{equation}
  \label{eq:6}
  w=r_{i_1}\cdot r_{i_2}\dots r_{i_k} \leftrightarrow (i_1, i_2,\dots,i_k)
\end{equation}
Then we will implement action of Weyl group elements on $\mathfrak{h}^*$ using the recipies from \cite{stembridge2001computational}, \cite{casselman1994machine}, but it will be done in the future. For now we are not implementing abstract construction of Weyl group since it is enough to be able to construct Weyl group orbits of $\mathfrak{h}^*$ elements.

Also sometimes we need to know the determinant of Weyl group element and we can take it into account using the following version of the {\tt{}reflect} function, called {\tt{}reflect-with-eps}.

\nwenddocs{}\nwbegincode{13}\sublabel{NWLitE-FunV-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-FunV-1}}}\moddef{Functions for root systems .scm~{\nwtagstyle{}\subpageref{NWLitE-FunV-1}}}\endmoddef
(define (co-root r)
  (div (mul 2 r) (prod r r)))

(define (reflect weight root)
  (sub weight (mul (prod weight (co-root root)) root)))

(define (reflect-with-eps weight root)
  (if (or (null? (cdr weight))
          (pair? (cdr weight)))
      (let ((v (reflect weight root)))
        (cons v (if (equal? weight v) 1 -1)))
      (let ((v (reflect (car weight) root)))
        (cons v (if (equal? (car weight) v) (cdr weight) (- (cdr weight)))))))

(define (eps weight)
  (if (or (null? (cdr weight))
          (pair? (cdr weight)))
      1
      (cdr weight)))

(define (no-eps weight)
  (if (or (null? (cdr weight))
          (pair? (cdr weight)))
      weight
      (car weight)))
\nwused{\\{NWLitE-lieF-1}}\nwendcode{}\nwbegindocs{14}\nwdocspar
Here {\it co-root} $\alpha^{\vee}$ is the element of the dual space $(\mathfrak{h}^*)^*$ which is identified with $\mathfrak{h}^*\approx \mathfrak{h}$ since we have scalar product on $\mathfrak{h}$.

Now we can implement wide range of root system related methods of {\tt{}semisimple-lie-algebra} class
\nwenddocs{}\nwbegincode{15}\sublabel{NWLitE-RooO-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-RooO-1}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NWLitE-RooO-1}}}\endmoddef
`(simple-co-roots ,(lambda (self)
              (map co-root (send self 'simple-roots))))
\nwalsodefined{\\{NWLitE-RooO-2}\\{NWLitE-RooO-3}\\{NWLitE-RooO-4}\\{NWLitE-RooO-5}\\{NWLitE-RooO-6}}\nwused{\\{NWLitE-SemZ-1}}\nwendcode{}\nwbegindocs{16}\nwdocspar
\begin{Def}
  {\bf Cartan matrix} with the elements
  \begin{equation}
    \label{eq:7}
    a_{ij}=(\alpha_i,\alpha_j^{\vee})=\frac{2(\alpha_i,\alpha_j)}{(\alpha_j,\alpha_j)}
  \end{equation}
  completely determines the set of simple roots and is useful for the classification and compact description of Lie algebras.
\end{Def}
\nwenddocs{}\nwbegincode{17}\sublabel{NWLitE-RooO-2}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-RooO-2}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NWLitE-RooO-1}}}\plusendmoddef
`(cartan-matrix
  ,(lambda (self)
     (map (lambda (a)
            (map (lambda (av) (prod a av)) (send self 'simple-co-roots)))
         (send self 'simple-roots))))
\nwendcode{}\nwbegindocs{18}\nwdocspar
We have already introduced Weyl group of reflections \ref{weyl-group-def}, now we implement procedure which constructs the union of Weyl group orbits of the set of weights.
\nwenddocs{}\nwbegincode{19}\sublabel{NWLitE-RooO-3}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-RooO-3}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NWLitE-RooO-1}}}\plusendmoddef
`(general-orbit , (lambda (self weights reflect)
    (let ((addon 
           (filter (lambda (x) (not (element-of-set? x weights)))
                   (fold-right union-set '() 
                               (map (lambda (w)
                                      (map
                                       (lambda (x) (reflect w x))
                                       (send self 'simple-roots)))
                                    weights)))))
      (if (null? addon)
          weights
          (send self 'general-orbit (union-set weights addon) reflect )))))

`(orbit ,
  (lambda (self weights)
    (send self 'general-orbit weights reflect)))

`(orbit-with-eps ,
  (lambda (self weights)
    (send self 'general-orbit weights reflect-with-eps)))
\nwendcode{}\nwbegindocs{20}\nwdocspar
\begin{Def}
  {\bf Fundamental weights} form a basis $w_1,\dots,w_r$ dual to the basis of simple co-roots $\alpha_1^{\vee},\dots,\alpha_r^{\vee}$.
\end{Def}
\nwenddocs{}\nwbegincode{21}\sublabel{NWLitE-RooO-4}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-RooO-4}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NWLitE-RooO-1}}}\plusendmoddef
`(fundamental-weights
  ,(lambda (self)
     (map (lambda (x)
            (sum (map-n (lambda (y z) (mul y z)) x (send self 'simple-roots))))
          (matrix-inverse (send self 'cartan-matrix)))))
\nwendcode{}\nwbegindocs{22}\nwdocspar
{\bf Weyl vector} $\rho=\sum_i w_i=\frac{1}{2}\sum_{\alpha\in \Delta^{+}}\alpha$.
\nwenddocs{}\nwbegincode{23}\sublabel{NWLitE-RooO-5}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-RooO-5}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NWLitE-RooO-1}}}\plusendmoddef
`(rho ,(lambda (self)
         (sum (send self 'fundamental-weights))))

\nwendcode{}\nwbegindocs{24}\nwdocspar
We can construct the full set of Lie algebra roots using Weyl symmetry.
\nwenddocs{}\nwbegincode{25}\sublabel{NWLitE-RooO-6}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-RooO-6}}}\moddef{Root system methods .scm~{\nwtagstyle{}\subpageref{NWLitE-RooO-1}}}\plusendmoddef
`(roots ,(lambda (self)
           (send self 'orbit (send self 'simple-roots))))
\nwendcode{}\nwbegindocs{26}\nwdocspar

@@

For the finite-dimensional representations we introduce abstract class {\tt{}representation} and concrete class {\tt{}highest-weight-representation}. Then we can describe  tensor product of two highest weight representations as the object of class {\tt{}representation} (but not of class {\tt{}highest-weight-representation}.
\nwenddocs{}\nwbegincode{29}\sublabel{NWLitE-RepK-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWLitE-RepK-1}}}\moddef{Representations .scm~{\nwtagstyle{}\subpageref{NWLitE-RepK-1}}}\endmoddef
(class 'representation 'object
       `(dim ())
       `(multiplicity , (lambda (self weight)
                          (error "Class is abstract!")))
       '(lie-algebra ())
       )
\nwalsodefined{\\{NWLitE-RepK-2}}\nwused{\\{NWLitE-lieF-1}}\nwendcode{}\nwbegindocs{30}\nwdocspar
